On the reality of the quantum state once again

### 2021-12-14 Articles/2021-HarriganSpekkens/HarriganSpekkens.tex
HarriganSpekkens: abstract

\begin{abstract}
We introduce an approach to the foundations of physics that is more in line with the foundations of mathematics. The idea is to examine current theories and find a set of starting physical assumptions that are sufficient to rederive them, which is the physical equivalent of the reverse mathematics program. This allows us to identify physical starting points (as opposed to mathematical ones), compare their deductive power, and find those that can better serve as a foundation for our theories.
\end{abstract}

@@

Let us first characterize the extent of the discepancy. In HS, the space of epistemic states $E_{HS}$ is the set of probability distributions over pure states, which is an infinite dimensional space. This is a direct consequence of defining the epistemic states as a distribution $P(\lambda|P)$ that, for each preparation $P$, assigns a well defined probability to each ontological state $\lambda$.\footnote{TODO: HS seems to be very vague about probability and probability density.} In QM, mixed states $E_{Q}$ are in one-to-one correspondence with the interior of the Bloch sphere, a three dimensional manifold. Looking at the difference in dimension, the discrepancy is not minor.

Note that the space of pure states $\Lambda$ is the same in both cases, one state for each direction of spin. Also note that all epistemic states in both spaces can be reached by mixture of pure states. The difference is that, in the quantum case, we have an ``aliasing'': different mixtures will give the same mixed states. Since each classical probability distribution corresponds to a unique mixing of pure states, we have a map $\iota : E_{HS} \mapsto E_{Q}$ that allows us to go from HS epistemic states to density matrixes. This map is surjective but not injective. Therefore $E_{Q}$ can be seen as a set of equivalence classes of $E_{HS}$. The key question, then, is the following: is the map $\iota$ sufficiently well behaved that we can understand the epistemic content of $E_{Q}$ based on the epistemic content of $E_{HS}$? Or is it so hopelessy broken that there is no connection between the two?

>>>>

\begin{abstract}
The ontological model framework introduced by Harrigan and Spekkens has been widely used to categorize quantum interpretation and as a basis for influential results such as the PBR theorem. In this paper we show that the model has a fundamental flaw: the epistemic structure it implicitly assumes does not follow the one dictated by quantum mechanics. Namely, the map between the epistemic states of the model and quantum density matrices is not even an order homomorphism in terms of the information entropy, which means the epistemic content of mixed states is not mapped in a meaningful way. Moreover, by recasting the definitions of the model directly in terms of density matrices, we find that only $\psi$-epistemic interpretations are allowed by quantum mechanics, a result that is the opposite of the PBR theorem. We show that the cause of this mismatch is the effect of non-contextuality over mixed preparations that does not allow for a single decomposition of mixed states over pure states.
\end{abstract}

@@

Let us first characterize the extent of the discrepancy. In HS, the space of epistemic states $E_{HS}$ is the set of probability distributions over pure states, which is an infinite dimensional space. This is a direct consequence of defining the epistemic states as a distribution $p(\lambda|P)$ that, for each preparation $P$, assigns a well defined probability $p$ to each ontological state $\lambda$.\footnote{TODO: HS seems to be very vague about probability and probability density.} In QM, mixed states $E_{Q}$ are in one-to-one correspondence with the interior of the Bloch sphere, a three dimensional manifold. Looking at the difference in dimension, the discrepancy is not minor.

Note that the space of pure states $\Lambda$ is the same in both cases, one state for each direction of spin. Also note that all epistemic states in both spaces can be reached by mixture of pure states. The difference is that, in the quantum (XXX) case, we have an ``aliasing'': different mixtures will give the same mixed states. Since each classical probability distribution corresponds to a unique mixing of pure states, we have a map $\iota : E_{HS} \mapsto E_{Q}$ that allows us to go from HS epistemic states to density matrices. This map is surjective but not injective. Therefore $E_{Q}$ can be seen as a set of equivalence classes of $E_{HS}$. The key question, then, is the following: is the map $\iota$ sufficiently well behaved that we can understand the epistemic content of $E_{Q}$ based on the epistemic content of $E_{HS}$? Or is it so hopelessy broken that there is no connection between the two?
